---
title: "Module6_Assignment1"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library("tidyverse")
library("tidymodels")
library("cluster")
```

```{r,message=FALSE, warning=FALSE}
trucks <- read_csv("trucks.csv")
```

### Task 1

```{r}

ggplot(trucks, aes(Distance, Speeding)) +
  geom_point(alpha = 0.4) + theme_bw()
```
There are 2 distinct clusterings at distances around 50 and 175.

### Task 2

```{r}
trucks = trucks %>%
  select(-Driver_ID)

kmeans_recipe = recipe(~ Distance + Speeding, trucks) 

trucks_dummy = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy, trucks) 
trucks_cleaned = bake(trucks_dummy, trucks)
  
```

### Task 3

```{r}
set.seed(64)
clusts = 
  tibble(k = 1:2) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```
```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```
The clusters above are very similar to the clusters in task one, confirming the 2 distinct clusters.


```{r}
clusters = kmeans(trucks_cleaned, centers = 2)
trucks = augment(clusters, trucks)
```

### Task 4

```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

Cluster set at 4 seems to be the most appropriate as the other clusters seem to be splitting the data too fine.

### Task 5

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

the value of K that appears to be best is 4 where the graph stops descending and forms a "hockey puck."  


### Task 6

```{r}
set.seed(64)
clusts = 
  tibble(k = 1:4) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

Cluster number 4 seems to be appropriate in this scenario as it seems to clearly show differentiation between the various sets.

