---
title: "Woody_Module2_Assign1_final"
output: word_document
---
## Task 1
```{r Task1}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest)
air=airquality
summary(air)
```
Ozone is the likely "Y" variable

## Task 2
```{r task2}

air2<- drop_na(air, Ozone, Solar.R)

```

## Task 3

```{r task3}
ggpairs(air2)
```
```{r task3 pt2}
ggcorr(air2, label = TRUE)
```
Day is the least associated with the Ozone variable
Temp is the most strongly associated with the Ozone variable.

## Task 4

```{r}
ggplot(air2,aes(x=Temp,y=Ozone))+geom_point(alpha=0.1) + geom_smooth(method = "lm", color = "red") + theme_bw()
```

## Task 5

```{r}
Ozone_Model = recipe(Ozone ~ Temp, air2)
Ozone_Model
```
```{r}
lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 
```

```{r}
lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(Ozone_Model)

lm_fit = fit(lm_wflow, air2)

summary(lm_fit$fit$fit$fit)
```
```{r}
confint(lm_fit$fit$fit$fit)
```

```{r}
ggplot(air2,aes(x=Temp,y=Ozone))+geom_point(alpha=0.1) + geom_smooth(method = "lm", se=FALSE) + theme_bw()
```

## Task 7

```{r}
testdata = data.frame(Temp = c(80,90,100))
predict(lm_fit, new_data = testdata)

-147.6461 + 2.4391*80
```
```{r}
#Prediction intervals
temp_var = predict(lm_fit$fit$fit$fit, interval = "prediction") #accessing the fit object with the three $fit
new_df = cbind(air2, temp_var)

ggplot(new_df, aes(x = Temp, y = Ozone)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  geom_line(aes(y=lwr), color = "red", linetype = "dashed") +
  geom_line(aes(y=upr), color = "red", linetype = "dashed") +
  theme_bw()
```
## Task 8

**Assumption 1:** The predictor and response variable have a linear relationship  
As noted above, it seems "reasonable" to say that there is something of a linear relationship between these two variables, but there is definitely a nonlinear effect present.  

**Assumption 2:** Model errors (residuals) are independent  
Let's use the Durbin-Watson Test to examine independence of residuals. The dwtest function is from the lmtest package.  
```{r}
dwtest(lm_fit$fit$fit$fit)
```
We fail to reject the null hypothesis with a p-value greater than 0.05. This suggests that the residuals are likely independent.  

**Assumption 3:** Model residuals exhibit constant variance  
Examine a plot of residuals.  
```{r}
air3 = air2 %>% mutate(resid1 = lm_fit$fit$fit$fit$residuals) #add the model residuals to our data frame
ggplot(air3,aes(x=Temp,y=resid1)) + geom_point() + theme_bw()
```
A non-linear effect is present.

**Assumption 4:** Model residuals are Normally-distributed  
Examine a histogram of the residuals.  
```{r}
ggplot(air3,aes(x=resid1)) + geom_histogram() + theme_bw()
```


```{r}

```

